### 2023-02-25 3_cleaning_race_df.Rmd
completed cleaning

>>>>

```{r}
black_df <- race_df %>% select(1:3, contains("Black"))

black_df <- black_df %>% 
  mutate(Race = "Black") %>% 
 rename("Death Rate (AIDSVu)" = `Black Death Rate`,
        "Population" = Black_pop,
        "Death Cases" = "Black Death Cases",
        "New Diagnoses Rate" = "New Diagnoses Black Rate",
        "New Diagnoses Cases" = "New Diagnoses Black Cases",
        "Percentage Diagnosed Late" = `Percent of Black Persons Diagnosed Late`,
        "Number of Persons Diagnosed Late" = `Number of Black Persons Diagnosed Late`,
        "Percentage Received Care" = `Percent of Black Persons Living with HIV who Received HIV Care`,
        "Percentage Virally Suppressed" = `Percent of Black Persons Living with HIV who were Virally Suppressed`,
        "HIV Prevalence Rate" = "Black Rate",
        "Prevalent HIV Cases" = "Black Cases",
        "Expected Death Cases" = black_exp_deaths,
        "Ratio Observed/Expected" = ratio_OE_black,
        "Death Rate (estimated)" = black_death_rate)

hispanic_df <- race_df %>% select(1:3, contains("Hispanic"))

hispanic_df <- hispanic_df %>% 
  mutate(Race = "Hispanic") %>% 
 rename("Death Rate (AIDSVu)" = `Hispanic Death Rate`,
        "Population" = Hispanic_pop,
        "Death Cases" = "Hispanic Death Cases",
        "New Diagnoses Rate" = "New Diagnoses Hispanic Rate",
        "New Diagnoses Cases" = "New Diagnoses Hispanic Cases",
        "Percentage Diagnosed Late" = `Percent of Hispanic Persons Diagnosed Late`,
        "Number of Persons Diagnosed Late" = `Number of Hispanic Persons Diagnosed Late`,
        "Percentage Received Care" = `Percent of Hispanic Persons Living with HIV who Received HIV Care`,
        "Percentage Virally Suppressed" = `Percent of Hispanic Persons Living with HIV who were Virally Suppressed`,
        "HIV Prevalence Rate" = "Hispanic Rate",
        "Prevalent HIV Cases" = "Hispanic Cases",
        "Expected Death Cases" = hispanic_exp_deaths,
        "Ratio Observed/Expected" = ratio_OE_hispanic,
        "Death Rate (estimated)" = hispanic_death_rate)

white_df <- race_df %>% select(1:3, contains("White"))

white_df <- white_df %>% 
  mutate(Race = "White") %>% 
 rename("Death Rate (AIDSVu)" = `White Death Rate`,
        "Population" = White_pop,
        "Death Cases" = "White Death Cases",
        "New Diagnoses Rate" = "New Diagnoses White Rate",
        "New Diagnoses Cases" = "New Diagnoses White Cases",
        "Percentage Diagnosed Late" = `Percent of White Persons Diagnosed Late`,
        "Number of Persons Diagnosed Late" = `Number of White Persons Diagnosed Late`,
        "Percentage Received Care" = `Percent of White Persons Living with HIV who Received HIV Care`,
        "Percentage Virally Suppressed" = `Percent of White Persons Living with HIV who were Virally Suppressed`,
        "HIV Prevalence Rate" = "White Rate",
        "Prevalent HIV Cases" = "White Cases",
        "Expected Death Cases" = white_exp_deaths,
        "Ratio Observed/Expected" = ratio_OE_white,
        "Death Rate (estimated)" = white_death_rate)

multiple_df <- race_df %>% select(1:3, contains("Multiple"))

multiple_df <- multiple_df %>% 
  mutate(Race = "Multiple") %>% 
 rename("Death Rate (AIDSVu)" = "Multiple Race Death Rate",
        "Population" = Multiple_pop,
        "Death Cases" = "Multiple Race Death Cases",
        "New Diagnoses Rate" = "New Diagnoses Multiple Race Rate",
        "New Diagnoses Cases" = "New Diagnoses Multiple Race Cases",
        "Percentage Received Care" = `Percent of Multiple Race Persons Living with HIV who Received HIV Care`,
        "Percentage Virally Suppressed" = `Percent of Multiple Race Persons Living with HIV who were Virally Suppressed`,
        "HIV Prevalence Rate" = "Multiple Races Rate",
        "Prevalent HIV Cases" = "Multiple Races Cases",
        "Expected Death Cases" = multiple_exp_deaths,
        "Ratio Observed/Expected" = ratio_OE_multiple,
        "Death Rate (estimated)" = multiple_death_rate)

race_df <- list(black_df, hispanic_df, white_df, multiple_df)

race_df <- reduce(race_df, full_join)
```

```{r}
saveRDS(race_df, here("data/processed_data/race_df_cleaned.rds"))
```

### 2023-02-25 code/processing_code/1_cleaning_2020.Rmd
completed cleaning

summary(deaths_race)
```
```{r}
new_dx <- new_dx %>% 
  filter(!grepl("DC", `State Abbreviation`)) %>%
  filter(!grepl("PR", `State Abbreviation`)) %>% 
  select(1:3, contains(c("male", "Black", "White", "Hispanic")), -contains(c("and", "Stability")))

@@

new_dx <- new_dx %>% 
  select(1:3,8:13)
```

>>>>

summary(deaths_race)
```
```{r}
new_dx <- new_dx %>% 
  filter(!grepl("DC", `State Abbreviation`)) %>%
  filter(!grepl("PR", `State Abbreviation`)) %>% 
  select(1:3, contains(c("male", "Black", "White", "Hispanic", "Multiple")), -contains(c("and", "Stability")))

@@

new_dx <- new_dx %>% 
  select(-contains(c("male", "Year")))
```

### 2023-02-25 code/processing_code/2_cleaning_sex_df.Rmd
completed cleaning

```{r}
sex_df <- readRDS(here("data/processed_data/sex_df.rds"))
```
Let's find the HIV death rate by sex
```{r}
sex_df <- sex_df %>% 
  mutate(M_Rate = round(`Male Death Cases`/pop_male*1e5),
         F_Rate = round(`Female Death Cases`/pop_female*1e5))
```

@@

>>>>

```{r}
sex_df <- readRDS(here("data/processed_data/sex_df.rds"))
```
Let's find the HIV death rate by sex
```{r}
sex_df <- sex_df %>% 
  mutate(Male_Rate = round(`Male Death Cases`/pop_male*1e5),
         Female_Rate = round(`Female Death Cases`/pop_female*1e5)) %>% 
  select(!`Year`)
```

@@

```{r}
female_df <- sex_df %>% select(1:3, contains("female")) 
male_df <- sex_df %>% select(1:3, !contains("female"))

female_df <- female_df %>% 
  mutate(Sex = "Female") %>% 
 rename("Death Rate (AIDSVu)" = `Female Death Rate`,
        "Population" = pop_female,
        "Death Cases" = "Female Death Cases",
        "New Diagnoses Rate" = "New Diagnoses Female Rate",
        "New Diagnoses Cases" = "New Diagnoses Female Cases",
        "Percentage Received Care" = "Female_rec_care_percent",
        "Percentage Virally Suppressed" = "Female_viral_sup_percent",
        "HIV Prevalence Rate" = "Female Rate",
        "Prevalent HIV Cases" = "Female Cases",
        "Expected Death Cases" = female_exp_deaths,
        "Ratio Observed/Expected" = ratio_OE_female,
        "Death Rate (estimated)" = Female_Rate)

male_df <- male_df %>% 
  mutate(Sex = "Male") %>% 
 rename("Death Rate (AIDSVu)" = `Male Death Rate`,
        "Population" = pop_male,
        "Death Cases" = "Male Death Cases",
        "New Diagnoses Rate" = "New Diagnoses Male Rate",
        "New Diagnoses Cases" = "New Diagnoses Male Cases",
        "Percentage Received Care" = "Male_rec_care_percent",
        "Percentage Virally Suppressed" = "Male_viral_sup_percent",
        "HIV Prevalence Rate" = "Male Rate",
        "Prevalent HIV Cases" = "Male Cases",
        "Expected Death Cases" = male_exp_deaths,
        "Ratio Observed/Expected" = ratio_OE_male,
        "Death Rate (estimated)" = Male_Rate) %>% 
  select(!`State Rate`)

sex_df <- full_join(female_df, male_df)

```

```{r}
saveRDS(sex_df, here("data/processed_data/sex_df_cleaned.rds"))
```
